---
title: "R Notebook - NicheNet analysis"
output: html_notebook
---

load libraries
```{r, echo=FALSE}
library(gridBase)
library(ComplexHeatmap)
library(nichenetr)
library(Seurat)
library(tidyverse)
library(circlize)
library(stringr)
library(RColorBrewer)
library(cowplot)
library(ggpubr)
library(xlsx)
```

load function and data
```{r}
ligand_target_matrix = readRDS(url("https://zenodo.org/record/3260758/files/ligand_target_matrix.rds"))
lr_network = readRDS(url("https://zenodo.org/record/3260758/files/lr_network.rds"))
weighted_networks = readRDS(url("https://zenodo.org/record/3260758/files/weighted_networks.rds"))
weighted_networks_lr_sig = weighted_networks$lr_sig %>% inner_join(lr_network %>% distinct(from,to), by = c("from","to"))
weighted_networks_lr_gr = weighted_networks$gr %>% inner_join(lr_network %>% distinct(from,to), by = c("from","to"))
##################
read_DEG<-function(loc_ranked_genes){
  py_clu_ma_f<-list.files(loc_ranked_genes,
                          full.names = T,pattern = ".xls")
  py_clu_ma<-lapply(py_clu_ma_f,function(x) read.xlsx(x,1))
  names(py_clu_ma)<-substr(py_clu_ma_f,nchar(loc_ranked_genes),nchar(py_clu_ma_f))
  names(py_clu_ma)
  #nrow(py_clu_ma[[4]])
  py_clu_ma
}
###################
read_csv_to_seutat<-function(loc_anndata){
  X<-as.matrix(read.table(paste0(loc_anndata,"/X.csv"),sep = ","))
  var<-read.table(paste0(loc_anndata,"/var.csv"),sep = ",",header = T)
  obs<-read.table(paste0(loc_anndata,"/obs.csv"),sep = ",",header = T)
  obsm<-read.table(paste0(loc_anndata,"/obsm.csv"),sep = ",",header = T)
  if (colnames(obs)[1]=="index") {
    colnames(X)<-as.character(var$index)
    rownames(obs)<-as.character(obs$index)
  }else{
    if (length(as.character(var$X))==0){
      colnames(X)<-rownames(var)
    }else{
      colnames(X)<-as.character(var$X)
    }
    rownames(obs)<-as.character(obs$X)
  }
  rownames(X)<-rownames(obs)
  umap_a<-as.matrix(obsm[,c(str_detect(colnames(obsm),"X_umap"))])
  rownames(umap_a)<-rownames(obs)
  colnames(umap_a)<-c("UMAP_1", "UMAP_2")
  
  seuratObj<-CreateSeuratObject(counts = t(X),
                     project = "1",
                     assay= "RNA")
  # I was not able to create the UMAP structure from scratch. This is why I used the structure of a older one which I overwrote.
  seuratObj@reductions$umap<-readRDS("I:/DATA/HPC/Server_Files/07_Scripts/reductions_umap.rds")
  seuratObj@reductions$umap@cell.embeddings<-umap_a
  seuratObj@meta.data<-obs
  if(length(which(colnames(seuratObj@meta.data) == "leiden_anno"))==1){
      pl1<-DimPlot(seuratObj, group.by = "leiden_anno")
      pl2<-DimPlot(seuratObj, group.by = "leiden")
      print(CombinePlots(list(pl1,pl2)))
  }else{
    if(length(which(colnames(seuratObj@meta.data) == "leiden"))==1){
      print(DimPlot(seuratObj, group.by = "leiden"))
    }
  }

  seuratObj
}
##############
combined_plotting_T<-function(p_ligand_pearson,
                            p_ligand_expression,
                            p_network,
                            p_tarec_expression,
                            seuObj,
                            mx_size,
                            fig.title){
  figures_without_legend = plot_grid(
    p_ligand_pearson + theme(legend.position = "none", axis.ticks = element_blank()) + theme(axis.title.x = element_text()) +
      ggtitle(fig.title),
    p_ligand_expression + theme(legend.position = "none", axis.ticks = element_blank()) + theme(axis.title.x = element_text()) + ylab(""),
    p_network + theme(legend.position = "none", axis.ticks = element_blank()) + ylab(""), 
    NULL,
    NULL,
    p_tarec_expression + theme(legend.position = "none", axis.ticks = element_blank()) + xlab(""), 
    align = "hv",
    nrow = 2,
    rel_widths = c(length(levels(seuObj)) * 2, length(levels(seuObj)) * 2, ncol(mx_size) ) ,
    rel_heights = c(nrow(mx_size), length(levels(seuObj)) + 5)) 
  
  legends = plot_grid(
    as_ggplot(get_legend(p_ligand_pearson)),
    as_ggplot(get_legend(p_ligand_expression)),
    as_ggplot(get_legend(p_network)),
    as_ggplot(get_legend(p_tarec_expression)),
    nrow = 2,
    align = "h")
  
  out_plot<-plot_grid(figures_without_legend, 
                      legends, 
                      rel_heights = c(10,2), nrow = 2, align = "hv")
  
  out_plot
}

combined_plotting_R<-function(p_ligand_expression,
                            p_network,
                            p_tarec_expression,
                            seuObj,
                            mx_size,
                            fig.title){
  figures_without_legend = plot_grid(
    p_ligand_expression + theme(legend.position = "none", axis.ticks = element_blank()) + theme(axis.title.x = element_text()) + ylab("") +
      ggtitle(fig.title),
    p_network + theme(legend.position = "none", axis.ticks = element_blank()) + ylab(""), 
    NULL,
    NULL,
    p_tarec_expression + theme(legend.position = "none", axis.ticks = element_blank()) + xlab(""), 
    align = "hv",
    nrow = 2,
    rel_widths = c(length(levels(seuObj)) * 4, ncol(mx_size) * 2 ) ,
    rel_heights = c(nrow(mx_size), length(levels(seuObj)) + 5)) 
  
  legends = plot_grid(
    as_ggplot(get_legend(p_ligand_expression)),
    as_ggplot(get_legend(p_network)),
    as_ggplot(get_legend(p_tarec_expression)),
    nrow = 2,
    align = "h")
  
  out_plot<-plot_grid(figures_without_legend, 
                      legends, 
                      rel_heights = c(10,2), nrow = 2, align = "hv")
  
  out_plot
}
```


read in DEGs and scrna dataset
```{r}
loc_ranked_genes<-"path DEG file"
loc_anndata<-"path to folder of dataser saved as .csv files"
save_path=loc_ranked_genes
listOfGenes<-read_DEG(loc_ranked_genes)
names(listOfGenes)<-substr(names(listOfGenes),12,nchar(names(listOfGenes))-5)
names(listOfGenes)
my_seu<-read_csv_to_seutat(loc_anndata)
my_seu_1<-my_seu
my_seu <- ScaleData(object = my_seu) 
colnames(my_seu@meta.data)

Cluster_Sel_Ident<-"leiden_anno"

DimPlot(my_seu,group.by = "leiden_anno")
```

specify sender and receiver and select associated DEGs
one of the next two chunks is executed for the respective directionality 
sender = MC3  MC2
```{r}
sender_celltypes=c("pericytes_1","pericytes_2")
receiver=c("tubulogenic_ECs")

py_clu_ma <-listOfGenes[[1]]
nr_goi<-c(2,3)
geneset_oi <- c(py_clu_ma[[nr_goi[1]]],py_clu_ma[[nr_goi[2]]])
```

sender = EC18
```{r}
receiver=c("pericytes_1","pericytes_2")
sender_celltypes=c("tubulogenic_ECs")

py_clu_ma <-listOfGenes[[1]]
nr_goi<-7
geneset_oi <- py_clu_ma[[nr_goi]]
```


preparation
```{r}
Idents(my_seu)<-"batch"
Idents(my_seu)<-"leiden_anno"
levels(my_seu)
levels(my_seu$leiden_anno)
# sender_celltypes=c("MC-2")
# receiver=c("EC-plexus_18")
condition_oi=c("d18m2", "d18a2")
write_textfile=T

lig_cut_off<-40
levels(my_seu)
seuratObj<-my_seu#subset(my_seu,idents=c("endothelial","mesenchymal"))


#which(str_detect(names(py_clu_ma),receiver)==T)

#geneset_oi <-as.character(py_clu_ma[[nr_goi]][,2])
geneset_oi <- geneset_oi %>% .[. %in% rownames(ligand_target_matrix)] # %>% 
geneset_oi
#.[. %in% as.character(Hum_Prot_atl_seq$Gene)]
print(paste0("geneset_oi: ",length(geneset_oi)))
gs_n<-names(py_clu_ma)[nr_goi]
print(gs_n)
fig.title<-paste0('c_oi_',paste(condition_oi,collapse = "_"),
                  '_gene_set_',paste(gs_n,collapse = "+"),
                  "_sender_", paste(sender_celltypes,collapse = "+"),
                  "_receiver_",paste(receiver,collapse = "+")
                  )
print(fig.title)


```

run
```{r}
print(paste0("receiver: ",receiver))
  #if ("batch" %in%  colnames(seuratObj@meta.data)) {
  Idents(seuratObj)<-'batch'
  seuratObj_coi<-subset(seuratObj, idents = condition_oi)
  
  # i=1
  # receiver<-receiver[i]
  # condition_oi<-condition_oi[i]
  # sender_celltypes<-sender_celltypes[i]
  #seuratObj_coi@meta.data
  Idents(seuratObj)<-'leiden_anno'
  Idents(seuratObj_coi)<-'leiden_anno'
  DimPlot(seuratObj_coi)
  levels(seuratObj_coi)
  levels(seuratObj)
  # new_Ids<-c("EC","EC","EC","epi","Mes","Mes")
  # names(new_Ids)<-levels(seuratObj)
  # seuratObj<-RenameIdents(seuratObj,new_Ids)
  expressed_genes_receiver = get_expressed_genes(receiver, seuratObj_coi, pct = 0.10)
  
  # expressed_genes_total<-as.character(raw_genes$index[raw_genes$n_cells>summary(raw_genes$n_cells)[2]])
  # expressed_genes_total<-unique(c(expressed_genes_total,expressed_genes_receiver))
  background_expressed_genes = expressed_genes_receiver %>% .[. %in% rownames(ligand_target_matrix)]
  #background_expressed_genes = expressed_genes_receiver %>% .[. %in% rownames(ligand_target_matrix)]
  print(paste0("Background Genes: ",length(background_expressed_genes)))
  
  # Idents(seuratObj_coi)<-'leiden'
  # sender_celltypes<-levels(seuratObj_coi)[1:5]
  #expressed_genes_sender = get_expressed_genes(sender_celltypes[2], seuratObj_coi, pct = 0.10)
  ## sender
  list_expressed_genes_sender = sender_celltypes %>% unique() %>% 
    lapply(get_expressed_genes, seuratObj_coi, 0.10) # lapply to get the expressed genes of every sender cell type separately here
  expressed_genes_sender = list_expressed_genes_sender %>% unlist() %>% unique()
  
  
  seurat_obj_receiver= subset(seuratObj, idents = receiver)
  # seurat_obj_receiver = SetIdent(seurat_obj_receiver, value = seurat_obj_receiver[["batch"]])
  Idents(seurat_obj_receiver)<-'leiden_anno'
  levels(seurat_obj_receiver)
  DimPlot(seurat_obj_receiver, reduction = "umap", group.by = "leiden_anno")
  # seurat_obj_receiver@meta.data[c("leiden_anno","batch")] %>% table()
  
  ## 3\. Define a set of potential ligands: these are ligands that are expressed by the “sender/niche” cell population and bind a (putative) receptor expressed by the “receiver/target” population
  ligands = lr_network %>% pull(from) %>% unique()
  receptors = lr_network %>% pull(to) %>% unique()
  
  expressed_ligands = intersect(ligands,expressed_genes_sender)
  expressed_receptors = intersect(receptors,expressed_genes_receiver)
  
  potential_ligands = lr_network %>% filter(from %in% expressed_ligands & to %in% expressed_receptors) %>% pull(from) %>% unique()
  
  
  # 4\) Perform NicheNet ligand activity analysis: rank the potential ligands based on the presence of their target genes in the gene set of interest (compared to the background set of genes)
  
  ligand_activities = predict_ligand_activities(geneset = geneset_oi, 
                                                background_expressed_genes = background_expressed_genes, 
                                                ligand_target_matrix = ligand_target_matrix, 
                                                potential_ligands = potential_ligands)
  
  ########################################################################
  #ligand_activities<-ligand_activities[as.character(ligand_activities$test_ligand) %in% rownames(seuratObj),]
  suspicious_gene<-"HLA"
  if (length(ligand_activities$test_ligand[str_detect(ligand_activities$test_ligand,suspicious_gene)])<1) {
    print("out")
    #  print(rownames(vis_ligand_target))
  }else{
    print("in")
    ligand_activities<-ligand_activities[-c(which(str_detect(ligand_activities$test_ligand,suspicious_gene)==T)),]
    # print(rownames(vis_ligand_target))
  }
  # ########################################################################

  ########################################################################
  
  ligand_activities = ligand_activities %>% arrange(-pearson) #%>% mutate(rank = rank(desc(pearson)))
  # show histogram of ligand activity scores
  p_hist_lig_activity = ggplot(ligand_activities, aes(x=pearson)) + 
    geom_histogram(color="black", fill="darkorange")  + 
    # geom_density(alpha=.1, fill="orange") +
    geom_vline(aes(xintercept=min(ligand_activities %>% top_n(lig_cut_off, pearson) %>% pull(pearson))), color="red", linetype="dashed", size=1) + 
    labs(x="ligand activity (PCC)", y = "# ligands") +
    theme_classic() + ggtitle(paste0('condition_oi: ',condition_oi,' | gene_set: ',gs_n," | sender: ",toString(sender_celltypes)," | receiver: ",receiver))
  p_hist_lig_activity
  #####################
  # seurat_obj_sender= subset(seuratObj, idents = sender_celltypes)
  #       ave_lig_exp<-AverageExpression(object = seurat_obj_sender,
  #                                              features = ligand_activities$test_ligand,
  #                                              slot = "scale.data")
  ######################
      ave_lig_exp<-AverageExpression(object = subset(seuratObj_coi,idents = sender_celltypes),
                                               features = ligand_activities$test_ligand,
                                               slot = "scale.data")
  colnames(ave_lig_exp$RNA)<-sender_celltypes
    ligand_activities<-cbind(ligand_activities, ave_lig_exp$RNA)
    ligand_activities<-ligand_activities[ligand_activities[sender_celltypes]>0,]
  
  
  best_upstream_ligands = ligand_activities %>% top_n(lig_cut_off, pearson) %>% #.[.[sender_celltypes] > 0]  %>% 
    arrange(-pearson) %>% pull(test_ligand) %>% unique()
  
    
    
    DotPlot(seuratObj_coi,group.by = 'leiden_anno', features = best_upstream_ligands %>% rev(), cols = "RdYlBu") + RotatedAxis()
    
    # 5\) Infer receptors and top-predicted target genes of ligands that are top-ranked in the ligand activity analysis
    
    ### Active target gene inference
    
    active_ligand_target_links_df = best_upstream_ligands %>%
      lapply(get_weighted_ligand_target_links,geneset = geneset_oi, ligand_target_matrix = ligand_target_matrix, n = 200) %>%
      bind_rows() %>% drop_na()
    
    active_ligand_target_links = prepare_ligand_target_visualization(ligand_target_df = active_ligand_target_links_df, ligand_target_matrix = ligand_target_matrix, cutoff = 0.33)
    
    order_ligands = intersect(best_upstream_ligands, colnames(active_ligand_target_links)) %>% rev() %>% make.names()
    order_targets = active_ligand_target_links_df$target %>% unique() %>% intersect(rownames(active_ligand_target_links)) %>% make.names()
    rownames(active_ligand_target_links) = rownames(active_ligand_target_links) %>% make.names() # make.names() for heatmap visualization of genes like H2-T23
    colnames(active_ligand_target_links) = colnames(active_ligand_target_links) %>% make.names() # make.names() for heatmap visualization of genes like H2-T23
    
    vis_ligand_target = active_ligand_target_links[order_targets,order_ligands] %>% t()
    colnames(vis_ligand_target)<-str_replace(colnames(vis_ligand_target),"[.]", "-")
    
    
    
    p_ligand_target_network = vis_ligand_target %>%
      make_heatmap_ggplot("Prioritized ligands","Predicted target genes",
                          color = "purple",legend_position = "top", x_axis_position = "top",
                          legend_title = "Regulatory potential")  + theme(axis.text.x = element_text(face = "italic")) +
      scale_fill_gradient2(low = "whitesmoke",  high = "purple"#, breaks = c(0,0.0025,0.005,0.01)
      )
    p_ligand_target_network
    
    
    #weighted_networks_lr_gr
    lr_network_top = lr_network %>% filter(from %in% best_upstream_ligands & to %in% expressed_receptors) %>% distinct(from,to)
    best_upstream_receptors = lr_network_top %>% pull(to) %>% unique()
    #######################
    lr_network_top_df_large = weighted_networks_lr_gr %>% filter(from %in% best_upstream_ligands & to %in% best_upstream_receptors)
    
    lr_network_top_df = lr_network_top_df_large %>% spread("from","weight",fill = 0)
    lr_network_top_matrix = lr_network_top_df %>% dplyr::select(-to) %>% as.matrix() %>% magrittr::set_rownames(lr_network_top_df$to)
    
    dist_receptors = dist(lr_network_top_matrix, method = "binary")
    
    ### Receptors of top-ranked ligands
    lr_network_top = lr_network %>% filter(from %in% best_upstream_ligands & to %in% expressed_receptors) %>% distinct(from,to)
    best_upstream_receptors = lr_network_top %>% pull(to) %>% unique()
    
    lr_network_top_df_large = weighted_networks_lr_sig %>% filter(from %in% best_upstream_ligands & to %in% best_upstream_receptors)
    
    lr_network_top_df = lr_network_top_df_large %>% spread("from","weight",fill = 0)
    lr_network_top_matrix = lr_network_top_df %>% dplyr::select(-to) %>% as.matrix() %>% magrittr::set_rownames(lr_network_top_df$to)
    
    dist_receptors = dist(lr_network_top_matrix, method = "binary")
    hclust_receptors = hclust(dist_receptors, method = "ward.D2")
    order_receptors = hclust_receptors$labels[hclust_receptors$order]
    
    dist_ligands = dist(lr_network_top_matrix %>% t(), method = "binary")
    hclust_ligands = hclust(dist_ligands, method = "ward.D2")
    order_ligands_receptor = hclust_ligands$labels[hclust_ligands$order]
    
    order_receptors = order_receptors %>% intersect(rownames(lr_network_top_matrix))
    order_ligands_receptor = order_ligands_receptor %>% intersect(colnames(lr_network_top_matrix))
    
    vis_ligand_receptor_network = lr_network_top_matrix[order_receptors, order_ligands_receptor]
    rownames(vis_ligand_receptor_network) = order_receptors %>% make.names()
    colnames(vis_ligand_receptor_network) = order_ligands_receptor %>% make.names()
    
    p_ligand_receptor_network_sig = vis_ligand_receptor_network %>% t() %>% 
      make_heatmap_ggplot("Ligands","Receptors", color = "mediumvioletred", 
                          x_axis_position = "top",legend_title = "Prior interaction potential")
    #
    ##
    #
    p_ligand_receptor_network_sig
    
    ### Receptors of top-ranked ligands, but after considering only bona fide ligand-receptor interactions documented in literature and publicly available databases
    lr_network_strict = lr_network %>% filter(database != "ppi_prediction_go" & database != "ppi_prediction")
    ligands_bona_fide = lr_network_strict %>% pull(from) %>% unique()
    receptors_bona_fide = lr_network_strict %>% pull(to) %>% unique()
    
    lr_network_top_df_large_strict = lr_network_top_df_large %>% distinct(from,to) %>% inner_join(lr_network_strict, by = c("from","to")) %>% distinct(from,to)
    lr_network_top_df_large_strict = lr_network_top_df_large_strict %>% inner_join(lr_network_top_df_large, by = c("from","to"))
    
    lr_network_top_df_strict = lr_network_top_df_large_strict %>% spread("from","weight",fill = 0)
    lr_network_top_matrix_strict = lr_network_top_df_strict %>% dplyr::select(-to) %>% as.matrix() %>% magrittr::set_rownames(lr_network_top_df_strict$to)
    
    dist_receptors = dist(lr_network_top_matrix_strict, method = "binary")
    hclust_receptors = hclust(dist_receptors, method = "ward.D2")
    order_receptors = hclust_receptors$labels[hclust_receptors$order]
    
    dist_ligands = dist(lr_network_top_matrix_strict %>% t(), method = "binary")
    hclust_ligands = hclust(dist_ligands, method = "ward.D2")
    order_ligands_receptor = hclust_ligands$labels[hclust_ligands$order]
    
    order_receptors = order_receptors %>% intersect(rownames(lr_network_top_matrix_strict))
    order_ligands_receptor = order_ligands_receptor %>% intersect(colnames(lr_network_top_matrix_strict))
    
    vis_ligand_receptor_network_strict = lr_network_top_matrix_strict[order_receptors, order_ligands_receptor]
    rownames(vis_ligand_receptor_network_strict) = order_receptors %>% make.names()
    colnames(vis_ligand_receptor_network_strict) = order_ligands_receptor %>% make.names()
    
    p_ligand_receptor_network_strict = vis_ligand_receptor_network_strict %>% t() %>% make_heatmap_ggplot("Ligands","Receptors", color = "mediumvioletred", x_axis_position = "top",legend_title = "Prior interaction potential\n(bona fide)")
    #
    ##
    #
    p_ligand_receptor_network_strict
    
    ## 6\) Summary visualizations of the NicheNet analysis
    
    
    
    # combined heatmap: overlay ligand activities with target genes
    ligand_pearson_matrix = ligand_activities %>% dplyr::select(pearson) %>% as.matrix() %>% magrittr::set_rownames(ligand_activities$test_ligand)
    
    rownames(ligand_pearson_matrix) = rownames(ligand_pearson_matrix) %>% make.names()
    colnames(ligand_pearson_matrix) = colnames(ligand_pearson_matrix) %>% make.names()
    
    vis_ligand_pearson = ligand_pearson_matrix[order_ligands, ] %>% as.matrix(ncol = 1) %>% magrittr::set_colnames("Pearson")
    # p_ligand_pearson = vis_ligand_pearson %>% make_heatmap_ggplot("Prioritized ligands","Ligand activity", color = "darkorange",legend_position = "top", x_axis_position = "top", legend_title = "Pearson correlation coefficient\n(target gene prediction ability)") + theme(legend.text = element_text(size = 9))
    
    p_ligand_pearson = vis_ligand_pearson %>% make_heatmap_ggplot("Prioritized ligands","Ligand activity", color = "darkorange",legend_position = "top", x_axis_position = "top", legend_title = "Pearson correlation coefficient\ntarget gene prediction ability)")
    p_ligand_pearson
    
    
    figures_without_legend = cowplot::plot_grid(p_ligand_pearson + theme(legend.position = "none", axis.ticks = element_blank()) +
                                                  ggtitle(paste0('condition_oi: ',condition_oi,' | gene_set: ',gs_n," | sender: ",toString(sender_celltypes)," | receiver: ",receiver)) +
                                                  theme(axis.title.x = element_text()),
                                                p_ligand_target_network + theme(legend.position = "none", axis.ticks = element_blank()) +
                                                  ylab(""),
                                                align = "hv",
                                                nrow = 1,
                                                rel_widths = c(ncol(vis_ligand_pearson)+10, ncol(vis_ligand_target)))
    
    legends = cowplot::plot_grid(
      ggpubr::as_ggplot(ggpubr::get_legend(p_ligand_pearson)),
      ggpubr::as_ggplot(ggpubr::get_legend(p_ligand_target_network)),
      nrow = 1,
      align = "h")
    
    
    combined_plot = cowplot::plot_grid(figures_without_legend, legends, rel_heights = c(10,2), nrow = 2, align = "hv")
    
    
    
    
    p1<-DimPlot(seurat_obj_receiver,group.by = "leiden_anno") #+ labs(title = "seurat_obj_receiver")
    p2<-DimPlot(seurat_obj_receiver,group.by = "leiden_anno") + labs(title = "seurat_obj_receiver")
    p3<-DimPlot(seuratObj_coi,group.by = Cluster_Sel_Ident)#,cols = scales::hue_pal()(2)[2]) #+ labs(title = "seuratObj_coi")
    p4<-DimPlot(seuratObj_coi,group.by = Cluster_Sel_Ident) + labs(title = "seuratObj_coi")
    p5<-DimPlot(seuratObj,group.by = Cluster_Sel_Ident) #+ labs(title = "seuratObj_coi")
    p6<-DimPlot(seuratObj,group.by = Cluster_Sel_Ident) + labs(title = "seuratObj")
    ######################################################################################################################################
    
    ######################################################################################################################################
    matrix_mean<-function(mm){
      mmax<-which(mm == max(mm), arr.ind = TRUE)
      mmin<-which(mm == min(mm), arr.ind = TRUE)
      vmax<-(mm[mmax[1,1],mmax[1,2]])
      vmin<-(mm[mmin[1,1],mmin[1,2]])
      (vmax+vmin)/2
    }
    ######################################################################################################################################
    com_plo_rli_exp<-function(tar_genes,seu_obj,y_achse){
      average_gene_expr_clu<-AverageExpression(object = subset(seu_obj,
                                                               idents = levels(seu_obj)),
                                               features = tar_genes,
                                               slot = "scale.data")
      vis_ligand_tumor_expression<-as.matrix(average_gene_expr_clu$RNA)
      color = colorRampPalette(rev(c("orangered4", "orangered", "gray90", "dodgerblue",
                                     "dodgerblue4")))(n = 50)
      #colnames(vis_ligand_tumor_expression)<-paste0("Clu_",colnames(vis_ligand_tumor_expression))
      x<-colnames(vis_ligand_tumor_expression)
      x<-gsub(" ", "_", x, fixed = TRUE)
      colnames(vis_ligand_tumor_expression)<-gsub("-", "_", x, fixed = TRUE)
      p_ligand_tumor_expression = vis_ligand_tumor_expression %>% 
        make_threecolor_heatmap_ggplot(y_achse,
                                       "Cell types", low_color = head(color)[1],
                                       mid_color = color[length(color) %/% 2], mid = 0,#matrix_mean(vis_ligand_tumor_expression),
                                       high_color = tail(color)[1],
                                       legend_position = "top", x_axis_position = "top",
                                       legend_title = "Expression\n(averaged over\nsingle cells)") + 
        theme(axis.text.y = element_text(face = "italic"))
      p_ligand_tumor_expression
    }
    
    
    com_plo_tarec_exp<-function(tar_genes,seu_obj,y_achse){
      average_gene_expr_clu<-AverageExpression(object = subset(seu_obj,
                                                               idents = levels(seu_obj)),
                                               features = tar_genes,
                                               slot = "scale.data") #"counts")#"
      vis_ligand_tumor_expression<-average_gene_expr_clu$RNA
      color = color = colorRampPalette(rev(brewer.pal(n = 7, name ="RdYlBu")))(100)
      
      matrix_mean(vis_ligand_tumor_expression)
      p_ligand_tumor_expression = t(vis_ligand_tumor_expression) %>% 
        make_threecolor_heatmap_ggplot(y_achse,
                                       "Cell types", low_color = head(color)[1],
                                       mid_color = color[length(color) %/% 2], mid = 0,#matrix_mean(vis_ligand_tumor_expression),
                                       high_color = tail(color)[1],
                                       legend_position = "top", x_axis_position = "top",
                                       legend_title = "Expression\n(averaged over\nsingle cells)") + 
        theme(axis.text.y = element_text(face = "italic"))
      p_ligand_tumor_expression
    }
    #######################################################################################################################################
    
    ######################################################################################################################################
    # Idents(seuratObj)<-"batch"
    SOTC<-my_seu#seuratObj#subset(seuratObj,idents=c(sender_celltypes,receiver))
    Idents(SOTC)<-Cluster_Sel_Ident
    # SOTC <- subset(SOTC,idents=c(sender_celltypes,receiver))
    # SOTC <- ScaleData(object = SOTC)
    co_bo_plo_tg<-combined_plotting_T(p_ligand_pearson = p_ligand_pearson,
                                      p_ligand_expression = com_plo_rli_exp(rownames(vis_ligand_pearson),SOTC,"Target ligands"),
                                      p_network = p_ligand_target_network,
                                      p_tarec_expression = com_plo_tarec_exp(colnames(vis_ligand_target),SOTC,"Clu"),
                                      SOTC,
                                      mx_size = vis_ligand_target,
                                      fig.title
    )
    co_bo_plo_tg
    
    rec_comb<-t(vis_ligand_receptor_network)
    rec_comb
    rec_comb <- rec_comb[match(rownames(vis_ligand_pearson), rownames(rec_comb)), ]
    
    p_ligand_receptor_comb = rec_comb  %>% #t() %>%
  make_heatmap_ggplot("Ligands","Receptors", color = "mediumvioletred", x_axis_position = "top",
                      legend_title = "Prior interaction potential\n(bona fide)")
    
    co_bo_plo_rec_li<-combined_plotting_T(p_ligand_pearson = p_ligand_pearson,
                                      p_ligand_expression = com_plo_rli_exp(rownames(rec_comb),SOTC,"Target ligands"),
                                      p_network = p_ligand_receptor_comb,
                                      p_tarec_expression = com_plo_tarec_exp(colnames(rec_comb),SOTC,"Clu"),
                                      SOTC,
                                      mx_size = rec_comb,
                                      fig.title
    )
    co_bo_plo_rec_li

    co_bo_plo_rec<-combined_plotting_R(p_ligand_expression = com_plo_rli_exp(colnames(vis_ligand_receptor_network_strict),
                                                                             SOTC,"Receptor ligands"),
                                       p_network = p_ligand_receptor_network_strict,
                                       p_tarec_expression = com_plo_tarec_exp(rownames(vis_ligand_receptor_network_strict),
                                                                              SOTC,"Clu"),
                                       SOTC,
                                       mx_size = t(vis_ligand_receptor_network_strict),
                                       fig.title
    )
    co_bo_plo_rec

    co_bo_plo_rec_f<-combined_plotting_R(p_ligand_expression = com_plo_rli_exp(colnames(vis_ligand_receptor_network),
                                                                               SOTC,"Receptor ligands"),
                                         p_network = p_ligand_receptor_network_sig,
                                         p_tarec_expression = com_plo_tarec_exp(rownames(vis_ligand_receptor_network),
                                                                                SOTC,"Clu"),
                                         SOTC,
                                         mx_size = t(vis_ligand_receptor_network),
                                         fig.title
    )
    co_bo_plo_rec_f
    ############################### to pdf ################################################################################################
    require(lattice)
    pdf(sprintf("%s/%s_%s_combined_plots.pdf", save_path,Sys.Date(),fig.title), width = 18, height = 12)
    # plot.new()
    print(p_hist_lig_activity)
    # print(DotPlot(seuratObj, features = best_upstream_ligands %>% rev(), cols = "RdYlBu") + RotatedAxis())
    # mtext(fig.title, outer=TRUE,  cex=1, line=-1)
    print(CombinePlots(list(p2,p4,p5,p6),ncol = 2))
    #print(DotPlot(seuratObj_coi, features = best_upstream_ligands %>% rev(), cols = "RdYlBu") + RotatedAxis())
    
    # std plots
    # print(p_ligand_target_network)
    # print(p_ligand_receptor_network_gr)
    # print(p_ligand_receptor_network_strict)
    # print(combined_plot)
    
    # plots with expression
    #print(co_bo_plo_rec_f)
    print(co_bo_plo_rec)
    print(co_bo_plo_rec_li)
    print(co_bo_plo_tg)
    
    ##vln##
    
    dev.off()
    
 ############# to TXT ################    
    if(write_textfile==T)
    {
        write.table(ligand_activities,
            sprintf("%s/%s_%s_ligand_activities.txt", save_path,Sys.Date(),fig.title),
            quote = F, sep = "\t")
        write.table(vis_ligand_pearson,
            sprintf("%s/%s_%s_vis_ligand_pearson.txt", save_path,Sys.Date(),fig.title),
            quote = F, sep = "\t")
        write.table(t(vis_ligand_receptor_network),
            sprintf("%s/%s_%s_receptor.txt", save_path,Sys.Date(),fig.title),
            quote = F, sep = "\t")
        write.table(t(vis_ligand_receptor_network_strict),
                    sprintf("%s/%s_%s_receptor_strict.txt", save_path,Sys.Date(),fig.title),
                    quote = F, sep = "\t")
        write.table(vis_ligand_target,
                    sprintf("%s/%s_%s_target.txt", save_path,Sys.Date(),fig.title),
                    quote = F, sep = "\t")
    }
```


selected ligand receptor pairs and interaction were composed for the cicro plot
```{r}
###########################
#######CICRO_Start#########
###########################
selected_reli_df<-read.xlsx(paste0(save_path,"/","Fig6_selected_lig_rec.xlsx"),1)

ligands_to_keep <- as.vector(as.matrix(selected_reli_df[c(3,5,7)]))
ligands_to_keep <- ligands_to_keep[!is.na(ligands_to_keep)]

col_MC_sender <- c("#0b8e00","#005e33","#64bc6f")
col_MC_sender <- c("purple","red","blue")
col_MC_sender <- c("black","red","blue")
lig_ty_na<-names(col_MC_sender)<-colnames(selected_reli_df[c(3,5,7)])
grid_col_ligand <- col_MC_sender


col_EC_receptors <- "#fb9a99"

scales::show_col(col_MC_sender)
print(col_MC_sender)
```


```{r}

ligand_type_indication_df = tibble(
    ligand_type = ligands_to_keep,
    ligand = ligands_to_keep)

for (i in lig_ty_na) {
  print(i)
  d<-selected_reli_df[[i]]
  d<-match(d[!is.na(d)],ligand_type_indication_df$ligand)
  d<-d[!is.na(d)]
  ligand_type_indication_df$ligand_type[d]<-rep(i,length(d))
}

    ### Infer target genes of top-ranked ligands and visualize in a circos plot

    active_ligand_target_links_df = best_upstream_ligands %>% lapply(get_weighted_ligand_target_links,geneset = geneset_oi, ligand_target_matrix = ligand_target_matrix, n = 250) %>% bind_rows()
    active_ligand_target_links_df = na.omit(active_ligand_target_links_df)

    active_ligand_target_links_df = active_ligand_target_links_df %>% mutate(target_type = "endothelial") %>% inner_join(ligand_type_indication_df) # if you want ot make circos plots for multiple gene sets, combine the different data frames and differentiate which target belongs to which gene set via the target type
    
circos_links = active_ligand_target_links_df
```

```{r}
grid_col_tbl_ligand = tibble(ligand_type = grid_col_ligand %>% names(), color_ligand_type = grid_col_ligand)
grid_col_tbl_target = tibble(target_type = grid_col_target %>% names(), color_target_type = grid_col_target)

    circos_links = circos_links %>% mutate(ligand = paste(ligand," ")) # extra space: make a difference between a gene as ligand and a gene as target!
    circos_links = circos_links %>% inner_join(grid_col_tbl_ligand) %>% inner_join(grid_col_tbl_target)
    links_circle = circos_links %>% dplyr::select(ligand,target, weight)

    ligand_color = circos_links %>% distinct(ligand,color_ligand_type)
    grid_ligand_color = ligand_color$color_ligand_type %>% set_names(ligand_color$ligand)
    target_color = circos_links %>% distinct(target,color_target_type)
    grid_target_color = target_color$color_target_type %>% set_names(target_color$target)

    grid_col =c(grid_ligand_color,grid_target_color)

    # give the option that links in the circos plot will be transparant ~ ligand-target potential score
    transparency = circos_links %>% mutate(weight =(weight-min(weight))/(max(weight)-min(weight))) %>% mutate(transparency = 1-weight) %>% .$transparency

    ## order ligands and targets
    target_order = circos_links$target %>% unique()
    ligand_order = c(ligand_type_indication_df$ligand) %>% c(paste(.," ")) %>% intersect(circos_links$ligand)
    order = c(ligand_order,target_order)

    # define the gaps between the different segments

    width_same_cell_same_ligand_type = 0.5
    width_different_cell = 6
    width_ligand_target = 15
    width_same_cell_same_target_type = 0.5

    gaps = c()
    for (i in 1:length(lig_ty_na)) {
      print(lig_ty_na[i])
      print(nrow(circos_links %>% filter(ligand_type == lig_ty_na[i])))
      # if(length(circos_links$ligand_type == lig_ty_na[i])==1){
      #   gaps = c(gaps,
      #            rep(width_same_cell_same_ligand_type, times = (circos_links %>%
      #                                                             filter(ligand_type == lig_ty_na[i]) %>% distinct(ligand) %>% nrow()))
      # }else{
      gaps = c(gaps,
               rep(width_same_cell_same_ligand_type, times = (circos_links %>%
                                                                filter(ligand_type == lig_ty_na[i]) %>% distinct(ligand) %>% nrow() -1))
      )
      #   print(gaps)
      # }
      if (i < length(lig_ty_na)){
        gaps = c(gaps,width_different_cell)
      }else{
        gaps = c(gaps,width_ligand_target)
      }

    }

    gaps = c(gaps,
             rep(width_same_cell_same_target_type, times = (circos_links %>% filter(target_type == "endothelial") %>% distinct(target) %>% nrow() -1)),
             width_ligand_target
    )
```


plotcircos
```{r}
circos.clear()

circos.par(gap.degree = gaps)
chordDiagram(links_circle, directional = 1,order=order,link.sort = TRUE, link.decreasing = FALSE, grid.col = grid_col,transparency = transparency, diffHeight = 0.005, direction.type = c("diffHeight", "arrows"),link.arr.type = "big.arrow", link.visible = links_circle$weight >= cutoff_include_all_ligands,annotationTrack = "grid",
             preAllocateTracks = list(track.height = 0.075))
title(paste0(fig.title,"_ligand_target"))
# we go back to the first track and customize sector labels
circos.track(track.index = 1, panel.fun = function(x, y) {
  circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index,
              facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.55), cex = 1)
}, bg.border = NA) #

```

Save circos plot to an svg file
```{r}
circos.clear()

svg(paste0(save_path,"/",Sys.Date(),"_",fig.title,"_","ligand_target_circos.svg"), width = 15, height = 18)
    circos.par(gap.degree = gaps)
    chordDiagram(links_circle, directional = 1,order=order,link.sort = TRUE, link.decreasing = FALSE, grid.col = grid_col,transparency = transparency, diffHeight = 0.005, direction.type = c("diffHeight", "arrows"),link.arr.type = "big.arrow", link.visible = links_circle$weight >= cutoff_include_all_ligands,annotationTrack = "grid",
                 preAllocateTracks = list(track.height = 0.075))
    title(paste0(fig.title,"_ligand_target"))
    # we go back to the first track and customize sector labels
    circos.track(track.index = 1, panel.fun = function(x, y) {
      circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index,
                  facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.55), cex = 1)
    }, bg.border = NA) #
    circos.clear()
    dev.off()
    ## png
    ##   2
```
